---
title: "Stat. 694 Project - Alameda Crime Data"
author: "Xiaodan Chen"
date: "2020/12/04"
output:
  ioslides_presentation: default
  slidy_presentation: default
  beamer_presentation: default
smaller: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Agenda

- Dataset

- Trends and Patterns

- Hypothesis Testing (Impact of COVID-19 on Crime Numbers)

- Heatmap in Cities

## Dataset

data.acgov.org - Alameda County Open Data Hub: 

<font size="4">https://data.acgov.org</font> 

<font size="2.5"> 

```{r include=FALSE}
library(pacman)
p_load(tidyverse, DT, dygraphs, plotly, lubridate, xts, ggmap, zoo, infer, fpp3)
crime <- read_csv('Crime_Reports.csv')
alam <- crime %>% select(DateTime, City, Longitude, Latitude, CrimeDescription, CrimeCode) %>%
  mutate(City = fct_collapse(City, 'HAYWARD' = c('HAYARD', 'HAWARD'), 
                             'SAN LEANDRO' = c('SAN LROENZO','SAN LOR','SAN LORENO','SAN  LORENZO', 'SAN LORENAZO', 'SAN  LEANDRO', 'SAN LORENZO', 'SAN LORNEZO'),
                             'CASTRO VALLEY' = c('CASTRO VALLE', 'CDASTRO VALLEY', 'CASRTO VALLEY', 'CASTRO', 'CATRO VALLEY'),
                             'LIVERMORE' = c('LIV', 'LIVEMORE', 'LIVERMORWE'),
                             'OAKLAND' = c('OAKALAND', 'OAKALND'),
                             'DUBLIN' = c('DUBLOIN', 'DUBLIIN')))

alam$time <- ymd_hms(alam$DateTime)
alam <- alam %>% mutate(year = year(time), month = as.yearmon(time), date = as.Date.POSIXct(time)) 

ala <- alam %>% 
  mutate(City=as.character(City)) %>%
  group_by(City) %>%
  summarize(n = n(), .groups = 'drop') %>%
  mutate(city_n = as.factor(ifelse(n > 800, City, 'OTHERS'))) %>%
  left_join(alam, City = City) 
```

```{r}
glimpse(alam)
```
</font>

## Trends and Patterns

```{r}
options(dplyr.summarise.inform = FALSE)
alame <- ala %>% filter(!is.na(month)) %>%
                    group_by(month) %>%
                    summarise(n_crime=n(), .groups = 'drop') %>%
                    mutate(city_n = 'TOTAL', city_n= as.factor(city_n))

alamed <- ala %>% filter(!is.na(month)) %>%
                    group_by(month, city_n) %>%
                    summarise(n_crime=n()) %>% 
                    mutate(city_n= as.factor(city_n)) %>%
                    rbind(alame)%>%
                    filter(month < 'June 2021')
plot1 <- alamed %>% ggplot(aes(month, n_crime)) +
  geom_line(aes(col = city_n)) +
  labs(col = 'city') +
  ylab('crime number')
  
ggplotly(plot1) 
```

## Trends and Patterns

```{r}
peak <- tibble(date = c(as.numeric(as.yearmon('2012-12-17')),
                        as.numeric(as.yearmon('2013-12-18')),
                        as.numeric(as.yearmon('2014-12-20')),
                        as.numeric(as.yearmon('2018-12-22')),
                        as.numeric(as.yearmon('2019-12-20'))))

plot2 <- plot1 + geom_vline(data = peak, aes(xintercept = date), 
                         linetype = 'dotted', col = 'blue4', size = 0.2) + 
  geom_text(data = peak, aes(x=date, y = c(2360, 2500, 2340, 2270, 2020),
                             label = c('2012-12-17', '2013-12-18', '2014-12-20', '2018-12-22', '2019-12-20')), col ='blue4')
ggplotly(plot2)
```

---

<div class="footer" style="margin-top:-40px;font-size:80%;">
<span style="color:black"><font size="6">Time Series for Different CrimeCode</font> </span>

<div class="footer" style="margin-top:30px;font-size:80%;">

<font size="3">999: 72 HOUR MENTAL HEALTH</font>    
<font size="3">90D: DUI ALCOHOL/DRUGS</font>   
<font size="3">35A: USE/UNDER INFLUENCE OF CONTROLLED SUBSTANCE</font>

```{r}
options(dplyr.summarise.inform = FALSE)
a <- alam %>% group_by(month, CrimeCode) %>%
  summarise(n_crime=n()) %>%
  arrange(desc(n_crime)) %>%
  filter(month < 'June 2021') %>%
  ggplot(aes(month, n_crime, col = CrimeCode)) +
  ylab('crime number') +
  geom_line() +
  geom_vline(data = peak, aes(xintercept = date), 
                         linetype = 'dotted', col = 'blue3', size =0.15) + 
  geom_text(data = peak, aes(x=date, y = c(750, 810, 540, 820, 720),
                             label = c('2012-12-17', '2013-12-18', '2014-12-20', '2018-12-22', '2019-12-20')), col ='blue4')
a <- ggplotly(a)
hide_legend(a)
```

## Hypothesis Testing 

<span style="color:black"><div class="footer" style="margin-top:80px;font-size:80%;">
<font size="5">Test 1.   

Impact of Lock Down on Alameda Crime Numbers.  
<div class="footer" style="margin-top:10px;font-size:80%;">
(Mar. 17, 2020 - May -22, 2020)

<div class="footer" style="margin-top:80px;font-size:80%;">
Test 2.    

Impact of COVID-19 on Alameda Crime Numbers.   
<div class="footer" style="margin-top:10px;font-size:80%;">
(2020)</font>  

---

<div class="footer" style="margin-top:-15px;font-size:80%;">
<span style="color:black"><font size="5.6">Test 1 - Impact of Lock Down on Alameda Crime Numbers</font>
<div class="footer" style="margin-top:20px;font-size:80%;">

```{r}
period <- tibble(start = c(as.numeric(as.yearmon('2018-03-17')),
                        as.numeric(as.yearmon('2019-03-17')),
                        as.numeric(as.yearmon('2020-03-17'))),
                  end = c(as.numeric(as.yearmon('2018-06-01')),
                        as.numeric(as.yearmon('2019-06-01')),
                        as.numeric(as.yearmon('2020-06-01'))),
                 lab1 = c('Mar. 17','Mar. 17','Mar. 17'),
                 lab2 = c('May 22, 2018','May 22, 2019','May 22, 2020'))

lockdown <- tibble(date = c(as.numeric(as.yearmon('2020-03-17')), 
                            as.numeric(as.yearmon('2020-05-22')),
                            as.numeric(as.yearmon('2018-03-17')), 
                            as.numeric(as.yearmon('2018-05-22')),
                            as.numeric(as.yearmon('2019-03-17')), 
                            as.numeric(as.yearmon('2019-05-22'))))
                           
plot1 %+% filter(alamed, month > 'Jan 2018') +
  geom_vline(data = lockdown, aes(xintercept = date), linetype = 'dotted', col = 'blue4', size = 0.4) +
  geom_text(aes(x = as.numeric(as.yearmon('2020-04-22')), y = 2050), label = 'Lock down', col = 'blue4', size = 3) +
  geom_text(data = period, aes(x = start, y =1130, label = lab1), col = 'blue4', size =2.5) +
  geom_text(data = period, aes(x = end, y =1030, label = lab2), col = 'blue4', size =2.5)
```

---

<div class="footer" style="margin-top:-25px;font-size:80%;">
<span style="color:black"><font size="5.6">Test 1 - Impact of Lock Down on Alameda Crime Numbers</font>
<div class="footer" style="margin-top:40px;font-size:80%;">
<font size="4">$H_0 : \mu = 53.369$</font>
<div class="footer" style="margin-top:3px;font-size:80%;">   
<font size="4">$H_1 : \mu \neq 53.369$</font>
<div class="footer" style="margin-top:20px;font-size:80%;">  
<font size="3"> $\mu = 53.369$   
mean crime number per day, March 17 - May 22, 2012 - 2019.</font>

<font size="3"> $\bar{x} = 39.119$    
mean crime number per day, March 17 - May 22 (lock down), 2020. </font>

<font size="3"> n = 67 </font>

<span style="color:black">

```{r include=FALSE}
population_1 <- alam %>% 
  filter(!is.na(date)) %>%
  filter(date >= '2012-03-17'& date <= '2012-05-22'|
         date >= '2013-03-17'& date <= '2013-05-22'|
         date >= '2014-03-17'& date <= '2014-05-22'|
         date >= '2015-03-17'& date <= '2015-05-22'|
         date >= '2016-03-17'& date <= '2016-05-22'|
         date >= '2017-03-17'& date <= '2017-05-22'|
         date >= '2018-03-17'& date <= '2018-05-22'|
         date >= '2019-03-17'& date <= '2019-05-22') %>% 
  group_by(date) %>%
  summarise(n_crime=n())
mu_test1 = mean(population_1$n_crime)

sample_1 <- alam %>% 
  filter(!is.na(date)) %>%
  filter(date >= '2020-03-17'& date <= '2020-05-22') %>% 
  group_by(date) %>%
  summarise(n_crime=n())
```

```{r}
t.test(sample_1$n_crime, mu = mu_test1)
```

---

<div class="footer" style="margin-top:-25px;font-size:80%;">
<span style="color:black"><font size="5.6">Test 2 - Impact of COVID-19 on Alameda Crime Numbers, 2020</font> 
<div class="footer" style="margin-top:40px;font-size:80%;">
<font size="4">$H_0 : \mu = 54.348$</font> 
<div class="footer" style="margin-top:3px;font-size:80%;">   
<font size="4">$H_1 : \mu \neq 54.348$</font>
<div class="footer" style="margin-top:20px;font-size:80%;">  
<font size="3"> $\mu = 54.348$   
mean crime number per day, 2012-01-01 - 2019-12-31.</font>

<font size="3"> $\bar{x} = 44.377$    
mean crime number per day, 2020-01-01 - 2020-11-27. </font>

<font size="3"> n = 334 </font>


```{r}
population_2 <- alam %>% 
  filter(!is.na(date)) %>%
  filter(date < '2020-01-01') %>% 
  group_by(date) %>%
  summarise(n_crime=n())
mu_test2 <- mean(population_2$n_crime)

sample_2 <- alam %>% 
  filter(!is.na(date)) %>%
  filter(date >= '2020-01-01') %>% 
  group_by(date) %>%
  summarise(n_crime=n())
```

```{r}
t.test(sample_2$n_crime, mu=mu_test2)
```

## Mapping

```{r message=FALSE}
library(ggmap)

# register_google(key = "AIzaSyCti8SYazuL9WnT6fkP_vheRWGrrkcfAcM", write = TRUE)

base <- get_map('Dublin, CA', zoom = 10, maptype = 'roadmap')

ggmap(base) + 
  geom_point(data = ala, aes(x=Longitude, y=Latitude, col=city_n), size =0.1) +
  labs(col = 'city') +
  ggtitle('Alameda County Crime Events 2012-2020')
```

## Heatmap in Cities

```{r}
alameda <- ala %>% 
  group_by(Longitude, Latitude) %>%
  summarise(loc_n=n()) %>%
  arrange(loc_n)

ggmap(base) + 
  geom_point(data = alameda, aes(x=Longitude, y=Latitude, col = loc_n)) +
  scale_colour_gradient(high="#132B43", low = "#F8766D") + 
  labs(col = 'crime number', size ='crime number') +
  ggtitle('Alameda County Crime Events 2012-2020')
```

## Heatmap in Hayward

```{r message=FALSE}
hayward <- ala %>% filter(City == 'HAYWARD',  year == '2020') %>%
  group_by(Longitude, Latitude) %>%
  summarise(crime_number=n()) %>%
  arrange(crime_number)

base_1 <- get_map('Hayward', zoom = 13, maptype = 'roadmap')

b <- ggmap(base_1) + 
  geom_point(data = hayward, aes(x=Longitude, y=Latitude, col = crime_number)) +
  labs(col = 'crime number', size ='crime number') +
  ggtitle('Hayward Crime Events 2020') +
  scale_colour_gradient(high="black", low = "#00BFC4") 

ggplotly(b)
```

## Hayward - CrimeCode

```{r message=FALSE}
hayward <- ala %>% filter(City == 'HAYWARD', month < 'Dec 2020' & month > 'Aug 2020') %>%
  group_by(Longitude, Latitude, CrimeCode) %>%
  summarise(crime_number=n()) 

hayward_1 <- hayward %>% filter(CrimeCode == '999')
hayward_2 <- hayward %>% filter(CrimeCode != '999')
  
c <- ggmap(base_1) + 
  geom_point(data = hayward_1, 
             aes(x=Longitude, y=Latitude, size = crime_number, col = CrimeCode)) +
  geom_point(data = hayward_2, 
             aes(x=Longitude, y=Latitude, size = crime_number, col = CrimeCode)) +
  ggtitle('Hayward Crime Events (September - November, 2020)') 
c <- ggplotly(c)
hide_legend(c)
```

## Datatable

<font size="2"> 

```{r warning=FALSE}
sample <- alam %>% sample_n(50)
datatable(sample)
```
</font>

---

<div class="centered" style="margin-top:280px;font-size:170%;">
<span style="color:black">Thanks!</font>



